State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-05-04 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
change to the experimental figure now it is plotted directly with subplot in Python to avoid spacing between plots in latex I must say it is a piece of art
change to the experimental graph again

Similar considerations need to be made regarding the representation of the dynamic field $v_t(\mathbf{r})$ using the basis function decomposition. The minimum distance between adjacent basis functions also needs to satisfy Shannon's sampling theorem. Thus, the minimum distance between basis must satisfy 
\begin{equation}
	\Delta_b < \frac{1}{2\nu_c}. 
\end{equation}
For the Gaussian basis functions $\phi(\mathbf{r})$ being considered, the basis width parameter $\sigma_{\phi}^2$ can also be inferred from spectral considerations \cite{Sanner1992,Scerri2009}. The Fourier transform of an \textit{n}-dimensional Gaussian is another Gaussian given by
\begin{equation}\label{eq:GaussianFT}
\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)^{\frac{n}{2}}\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right).
\end{equation}
where 
\begin{equation}\label{eq:GaussianFTWidth}
	\sigma^2_{\nu} = \frac{1}{\pi^2\sigma_{\phi}^2}. 
\end{equation}
width $\sigma^2_{\nu}$ should be chosen to give at least 3 dB attenuation at $\boldsymbol\nu_c$ giving
\begin{equation}\label{eq:WidthFrequencyRelationship}
 \sigma^2_{\nu}=\frac{2\boldsymbol\nu_c^\top \boldsymbol\nu_c}{\ln 2}.
\end{equation}
This ensures that the basis functions can represent field with frequency content up to $\boldsymbol\nu_c $. Proof of \ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} is given in \ref{ap:FrequencyAnalysis}.
% Thus to ensure that the basis functions can represent fields with frequency content up to $\boldsymbol{\nu}_c$, $\sigma_{\nu}$ should be set with at least the 3~dB frequency for the basis functions equal to $\boldsymbol{\nu}_c$, thus obtained $\sigma^2_{\nu}$ and therefore $\sigma_{\phi}^2$.

@@

Since the observation equation \ref{ObservationEquation} is linear the update step equations are simplified to the standard Kalman filtering formulations. The state estimates from the forward filtering are used to form a new set of sigma points for the smoother, as described above. However, the predicted cross-covariance of the states needs to be calculated. A summary of the URTSS procedure is given in algorithm~\ref{UKFAlgorithm}. It should be noted that as the disturbance, $e_t\left(\mathbf{r}\right)$ and the measurement noise $ \boldsymbol{\varepsilon}_t$ are additive the additive form of the URTS can be used. This reduces the dimension and the number of sigma points used in the algorithm.

@@

The sensors were spaced in a regular square grid, where the number of sensors was 100, sensor width $\sigma^2_m$ was 1.44 which equals $2$~mm full width at half maximum, with spacing $\Delta_y$ equal to $2.22$~mm. This arrangement modelled the recording as having some cross talk, which is typical from neurophysiological recordings. Note, sensor design to guarantee representation of the field can be achieved using the frequency analysis from section~\ref{SpectralAnalysisSection} given access to the true field. \omg{From frequency analysis the separation is $2.77$, with $\rho = 2$ (note same as basis function).} 
% \begin{figure}\label{fig:experimental design}
% 	\begin{center}
% 		\includegraphics[scale=0.6]{./Graph/ExperimentFigure.pdf} 
% 	\end{center}
% 	\caption{Example of experimental design and model output. \textbf{A}. Example of the neural field with sensors. The centres of the sensors are shown by crossed and the sensor widths ($1/2$ height of Gaussian) are shown by the circles. \textbf{B} Example of data generated by the model through the observations. \textbf{C}. Power spectral density of the data generated by an observation showing the typical $1/f$ characteristic of iEEG.} \label{fig:FrequencyAnalysis} 
% \end{figure}

\begin{figure}[!hb]
\centering
\subfigure[]{\includegraphics[width=0.3\textwidth]{./Graph/ExperimentFigurePy.pdf}}\\
% \hspace{2.5mm}
\subfigure[]{\includegraphics[width=0.3\textwidth]{./Graph/ExperimentFigureEEGPy.pdf}}\\
% \hspace{2.5mm}
\subfigure[]{\includegraphics[width=0.3\textwidth]{./Graph/ExperimentFigurePSDPy.pdf}}
\caption{Example of experimental design and model output. \textbf{A}. Example of the neural field with sensors. The centres of the sensors are shown by crossed and the sensor widths ($1/2$ height of Gaussian) are shown by the circles. \textbf{B} Example of data generated by the model through the observations. \textbf{C}. Power spectral density of the data generated by an observation showing the typical $1/f$ characteristic of iEEG.} \label{fig:experimental design} 
\end{figure}
\begin{figure}[th]
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/zeta.pdf}}
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/theta0.pdf}}\\
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/theta1.pdf}}
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/theta2.pdf}}\\
\caption{Distribution of the parameters estimates over 272
runs.}
\label{fig:Parameters}
\end{figure}
\subsection{Spatial Frequency Analysis} In accordance with Section~\ref{SpectralAnalysisSection}, the spatial frequency of the field was used to confirm the spacing and width of the sensors was adequate to capture the significant dynamics of the field. Figure~\ref{} shows the spatial frequency of the field. The cut-off frequency, $\nu_c$, was taken to be \omg{??}, which corresponds to the \omg{??-3~dB} from the peak power. From equation~\ref{eq:MinimumSensorDistance}, this yielded a minimum sensor separation of \omg{??}. This confirms our separation was enough to prevent problems associated with aliasing.

>>>>

Similar considerations need to be made regarding the representation of the dynamic field $v_t(\mathbf{r})$ using the basis function decomposition. The minimum distance between adjacent basis functions also needs to satisfy Shannon's sampling theorem. Thus, the minimum distance between basis must satisfy 
\begin{equation}
	\Delta_b < \frac{1}{2\nu_c}. 
\end{equation}
For the Gaussian basis functions $\phi(\mathbf{r})$ being considered, the basis width parameter $\sigma_{\phi}^2$ can also be inferred from spectral considerations \cite{Sanner1992,Scerri2009}. The Fourier transform of an \textit{n}-dimensional Gaussian is another Gaussian given by
\begin{equation}\label{eq:GaussianFT}
\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)^{\frac{n}{2}}\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right).
\end{equation}
where 
\begin{equation}\label{eq:GaussianFTWidth}
	\sigma^2_{\nu} = \frac{1}{\pi^2\sigma_{\phi}^2}. 
\end{equation}
width $\sigma^2_{\nu}$ should be chosen to give at least 3 dB attenuation at $\boldsymbol\nu_c$ giving
\begin{equation}\label{eq:WidthFrequencyRelationship}
 \sigma^2_{\nu}=\frac{2\boldsymbol\nu_c^\top \boldsymbol\nu_c}{\ln 2}.
\end{equation}
This ensures that the basis functions can represent field with frequency content up to $\boldsymbol\nu_c $. Proof of \ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} is given in Appendix~\ref{ap:FrequencyAnalysis}.
% Thus to ensure that the basis functions can represent fields with frequency content up to $\boldsymbol{\nu}_c$, $\sigma_{\nu}$ should be set with at least the 3~dB frequency for the basis functions equal to $\boldsymbol{\nu}_c$, thus obtained $\sigma^2_{\nu}$ and therefore $\sigma_{\phi}^2$.

@@

Since the observation equation \ref{ObservationEquation} is linear the standard Kalman filter can be used to update the state equations. The state estimates from the forward filtering are used to form a new set of sigma points for the smoother, as described above. In order to compute the smoother gain the predicted cross-covariance matrix of the states needs to be calculated. A summary of the URTSS procedure is given in algorithm~\ref{UKFAlgorithm}. It should be noted that as the disturbance, $e_t\left(\mathbf{r}\right)$ and the measurement noise $ \boldsymbol{\varepsilon}_t$ are additive the additive form of the URTS can be used. This reduces the dimension and the number of sigma points used in the algorithm.

@@

The sensors were spaced in a regular square grid, where the number of sensors was 100, sensor width $\sigma^2_m$ was 1.44 which equals $2$~mm full width at half maximum, with spacing $\Delta_y$ equal to $2.22$~mm. This arrangement modelled the recording as having some cross talk, which is typical from neurophysiological recordings. Note, sensor design to guarantee representation of the field can be achieved using the frequency analysis from section~\ref{SpectralAnalysisSection} given access to the true field. \omg{From frequency analysis the separation is $2.77$, with $\rho = 2$ (note same as basis function).} 
  \begin{figure}\label{fig:experimental design}
  	\begin{center}
  		\includegraphics[width=0.3\textwidth]{./Graph/ExperimentFigurePy_1.pdf} 
  	\end{center}
  	\caption{Example of experimental design and model output. \textbf{A}. Example of the neural field with sensors. The centres of the sensors are shown by crossed and the sensor widths ($1/2$ height of Gaussian) are shown by the circles. \textbf{B} Example of data generated by the model through the observations. \textbf{C}. Power spectral density of the data generated by an observation showing the typical $1/f$ characteristic of iEEG.} \label{fig:FrequencyAnalysis}
  \end{figure}

\begin{figure}[th]
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/zeta.pdf}}
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/theta0.pdf}}\\
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/theta1.pdf}}
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/theta2.pdf}}\\
\caption{Distribution of the parameters estimates over 272
runs.}
\label{fig:Parameters}
\end{figure}
\subsection{Spatial Frequency Analysis} In accordance with Section~\ref{SpectralAnalysisSection}, the spatial frequency of the field was used to confirm the spacing and width of the sensors was adequate to capture the significant dynamics of the field. Figure~\ref{} shows the spatial frequency of the field. The cut-off frequency, $\nu_c$, was taken to be \omg{??}, which corresponds to the \omg{??-3~dB} from the peak power. From equation~\ref{eq:MinimumSensorDistance}, this yielded a minimum sensor separation of \omg{??}. This confirms our separation was enough to prevent problems associated with aliasing.
